---
title: "Bookmark Modules Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bookmark-module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(shinystate)
```

## Introduction

The `shinystate` package was greatly inspired by an [example application](https://github.com/jcheng5/rpharma-demo) created by Joe Cheng (creator of Shiny) to accompany his keynote presentation at the 2018 [R/Pharma conference](https://rinpharma.com/). Among other notable features as documented in the GitHub repository [README](https://github.com/jcheng5/rpharma-demo/blob/master/README.md), the application provided an alternative user interface powered by Shiny modules to save and restore bookmarkable state. The following example is an adaptation of the original version to utilize `shinystate` to manage the bookmarkable state features.

## How to Run Application

The application source code is included in the 'shinystate' package and it can be launched with the following code:

```{r run-app-package, eval=FALSE}
library(shiny)
library(shinystate)
runExample("bookmark_module", package = "shinystate")
```

## Application Code

The remainder of this vignette shows the source code of the applicaiton. Note that in the example version included in the package, separate R scripts containing the module and utility function code are sourced in the main application.

### `app.R`

```{r main-app, eval=FALSE}
library(shiny)
library(shinystate)
library(dplyr)
library(DT)
library(rlang)
library(lubridate)

#  recommended to define a directory for storage or a pins board
storage <- StorageClass$new()

ui <- function(req) {
  tagList(
    # Bootstrap header
    tags$header(
      class = "navbar navbar-default navbar-static-top",
      tags$div(
        class = "container-fluid",
        tags$div(
          class = "navbar-header",
          tags$div(class = "navbar-brand", "Bookmark Module Demo")
        ),
        # Links for restoring/loading sessions
        tags$ul(
          class = "nav navbar-nav navbar-right",
          tags$li(
            bookmark_modal_load_ui("bookmark")
          ),
          tags$li(
            bookmark_modal_save_ui("bookmark")
          )
        )
      )
    ),
    fluidPage(
      use_shinystate(),
      sidebarLayout(
        position = "right",
        column(
          width = 4,
          wellPanel(
            select_vars_ui("select")
          ),
          wellPanel(
            filter_ui("filter")
          )
        ),
        mainPanel(
          tabsetPanel(
            id = "tabs",
            tabPanel("Plot", tags$br(), plotOutput("plot", height = 600)),
            tabPanel("Summary", tags$br(), verbatimTextOutput("summary")),
            tabPanel("Table", tags$br(), tableOutput("table"))
          )
        )
      )
    )
  )
}

server <- function(input, output, session) {
  callModule(bookmark_mod, "bookmark", storage)
  storage$register_metadata()
  datasetExpr <- reactive(expr(mtcars %>% mutate(cyl = factor(cyl))))
  filterExpr <- callModule(filter_mod, "filter", datasetExpr)
  selectExpr <- callModule(
    select_vars,
    "select",
    reactive(names(eval_clean(datasetExpr()))),
    filterExpr
  )

  data <- reactive({
    resultExpr <- selectExpr()
    df <- eval_clean(resultExpr)
    validate(need(nrow(df) > 0, "No data matches the filter"))
    df
  })

  output$table <- renderTable(
    {
      data()
    },
    rownames = TRUE
  )

  do_plot <- function() {
    plot(data())
  }

  output$plot <- renderPlot({
    do_plot()
  })

  output$summary <- renderPrint({
    summary(data())
  })

  output$code <- renderText({
    format_tidy_code(selectExpr())
  })
}

shinyApp(ui, server, onStart = function() {
  shiny::enableBookmarking("server")
})

```

### `bookmark_modules.R`

```{r bookmark-modules, eval=FALSE}
bookmark_modal_save_ui <- function(id) {
  ns <- NS(id)

  tagList(
    actionLink(ns("show_save_modal"), "Save session")
  )
}

bookmark_modal_load_ui <- function(id) {
  ns <- NS(id)

  tagList(
    actionLink(ns("show_load_modal"), "Restore session")
  )
}

bookmark_load_ui <- function(id) {
  ns <- NS(id)
  tagList(
    uiOutput(ns("saved_sessions"))
  )
}

bookmark_mod <- function(input, output, session, storage) {
  ns <- session$ns
  session_df <- reactive({
    storage$get_sessions()
  })

  output$saved_sessions_placeholder <- renderUI({
    DT::dataTableOutput(session$ns("saved_sessions_table"))
  })

  output$saved_sessions_table <- DT::renderDataTable({
    req(session_df())
    DT::datatable(
      session_df(),
      escape = FALSE,
      selection = "single"
    )
  })

  session_choice <- reactive({
    req(session_df())
    req(input$saved_sessions_table_rows_selected)
    i <- input$saved_sessions_table_rows_selected
    url <- session_df()[i, "url"]
    return(url)
  })

  observeEvent(input$restore, {
    req(session_choice())
    storage$restore(session_choice())
  })

  shiny::setBookmarkExclude(c(
    "show_save_modal",
    "show_load_modal",
    "save_name",
    "save",
    "session_choice",
    "restore"
  ))

  observeEvent(input$show_load_modal, {
    showModal(modalDialog(
      size = "xl",
      easyClose = TRUE,
      title = "Restore session",
      footer = tagList(
        modalButton("Cancel"),
        actionButton(session$ns("restore"), "Restore", class = "btn-primary")
      ),
      tagList(
        uiOutput(session$ns("saved_sessions_placeholder"))
      )
    ))
  })

  observeEvent(input$show_save_modal, {
    showModal(modalDialog(
      easyClose = TRUE,
      textInput(session$ns("save_name"), "Give this session a name"),
      footer = tagList(
        modalButton("Cancel"),
        actionButton(session$ns("save"), "Save", class = "btn-primary")
      )
    ))
  })

  observeEvent(input$save, ignoreInit = TRUE, {
    tryCatch(
      {
        if (!isTruthy(input$save_name)) {
          stop("Please specify a bookmark name")
        } else {
          removeModal()
          storage$snapshot(
            session_metadata = list(
              save_name = input$save_name,
              timestamp = Sys.time()
            )
          )
          showNotification(
            "Session successfully saved"
          )
        }
      },
      error = function(e) {
        showNotification(
          conditionMessage(e),
          type = "error"
        )
      }
    )
  })
}
```

### `filter_module.R`

```{r filter-module, eval=FALSE}
filter_ui <- function(id) {
  ns <- NS(id)

  tagList(
    div(id = ns("filter_container")),
    actionButton(ns("show_filter_dialog_btn"), "Add filter")
  )
}

filter_mod <- function(input, output, session, data_expr) {
  ns <- session$ns

  setBookmarkExclude(c("show_filter_dialog_btn", "add_filter_btn"))

  filter_fields <- list()
  makeReactiveBinding("filter_fields")

  onBookmark(function(state) {
    state$values$filter_field_names <- names(filter_fields)
  })

  onRestore(function(state) {
    filter_field_names <- state$values$filter_field_names
    for (fieldname in filter_field_names) {
      addFilter(fieldname)
    }
  })

  observeEvent(input$show_filter_dialog_btn, {
    available_fields <- names(eval_clean(data_expr())) %>%
      base::setdiff(names(filter_fields))

    showModal(modalDialog(
      title = "Add filter",

      radioButtons(ns("filter_field"), "Field to filter", available_fields),

      footer = tagList(
        modalButton("Cancel"),
        actionButton(ns("add_filter_btn"), "Add filter")
      )
    ))
  })

  observeEvent(input$add_filter_btn, {
    addFilter(input$filter_field)
    removeModal()
  })

  addFilter <- function(fieldname) {
    id <- paste0("filter__", fieldname)

    filter <- createFilter(
      data = eval_clean(data_expr())[[fieldname]],
      id = ns(id),
      fieldname = fieldname
    )

    freezeReactiveValue(input, id)

    insertUI(
      paste0("#", ns("filter_container")),
      "beforeEnd",
      # TODO: escape special characters in fieldname
      filter$ui
    )

    filter$inputId <- id
    filter_fields[[fieldname]] <<- filter
  }

  reactive({
    result_expr <- data_expr()

    if (length(filter_fields) == 0) {
      return(result_expr)
    }

    # Gather up all filter expressions
    exprs <- lapply(names(filter_fields), function(name) {
      filter <- filter_fields[[name]]
      x <- as.symbol(name) #df[[name]]
      param <- input[[filter[["inputId"]]]]
      cond_expr <- filter[["filterExpr"]](x = x, param = param)
      if (!is.null(cond_expr)) {
        result_expr <<- expr(!!result_expr %>% filter(!!cond_expr))
      }
      invisible()
    })

    result_expr
  })
}

createFilter <- function(data, id, fieldname) {
  UseMethod("createFilter")
}

createFilter.character <- function(data, id, fieldname) {
  list(
    ui = textInput(id, fieldname, ""),
    filterExpr = function(x, param) {
      if (!nzchar(param)) {
        NULL
      } else {
        expr(grepl(!!param, !!x, ignore.case = TRUE, fixed = TRUE))
      }
    }
  )
}

createFilter.numeric <- function(data, id, fieldname) {
  list(
    ui = sliderInput(
      id,
      fieldname,
      min = min(data),
      max = max(data),
      value = range(data)
    ),
    filterExpr = function(x, param) {
      expr(!!x >= !!param[1] & !!x <= !!param[2])
    }
  )
}

createFilter.integer <- createFilter.numeric

createFilter.factor <- function(data, id, fieldname) {
  inputControl <- if (length(levels(data)) > 6) {
    selectInput(id, fieldname, levels(data), character(0), multiple = TRUE)
  } else {
    checkboxGroupInput(id, fieldname, levels(data))
  }

  list(
    ui = inputControl,
    filterExpr = function(x, param) {
      if (length(param) == 0) {
        NULL
      } else {
        expr(!!x %in% !!param)
      }
    }
  )
}

createFilter.POSIXt <- createFilter.numeric
```

### `select_module.R`

```{r select-module, eval=FALSE}
filter_ui <- function(id) {
  ns <- NS(id)

  tagList(
    div(id = ns("filter_container")),
    actionButton(ns("show_filter_dialog_btn"), "Add filter")
  )
}

filter_mod <- function(input, output, session, data_expr) {
  ns <- session$ns

  setBookmarkExclude(c("show_filter_dialog_btn", "add_filter_btn"))

  filter_fields <- list()
  makeReactiveBinding("filter_fields")

  onBookmark(function(state) {
    state$values$filter_field_names <- names(filter_fields)
  })

  onRestore(function(state) {
    filter_field_names <- state$values$filter_field_names
    for (fieldname in filter_field_names) {
      addFilter(fieldname)
    }
  })

  observeEvent(input$show_filter_dialog_btn, {
    available_fields <- names(eval_clean(data_expr())) %>%
      base::setdiff(names(filter_fields))

    showModal(modalDialog(
      title = "Add filter",

      radioButtons(ns("filter_field"), "Field to filter", available_fields),

      footer = tagList(
        modalButton("Cancel"),
        actionButton(ns("add_filter_btn"), "Add filter")
      )
    ))
  })

  observeEvent(input$add_filter_btn, {
    addFilter(input$filter_field)
    removeModal()
  })

  addFilter <- function(fieldname) {
    id <- paste0("filter__", fieldname)

    filter <- createFilter(
      data = eval_clean(data_expr())[[fieldname]],
      id = ns(id),
      fieldname = fieldname
    )

    freezeReactiveValue(input, id)

    insertUI(
      paste0("#", ns("filter_container")),
      "beforeEnd",
      # TODO: escape special characters in fieldname
      filter$ui
    )

    filter$inputId <- id
    filter_fields[[fieldname]] <<- filter
  }

  reactive({
    result_expr <- data_expr()

    if (length(filter_fields) == 0) {
      return(result_expr)
    }

    # Gather up all filter expressions
    exprs <- lapply(names(filter_fields), function(name) {
      filter <- filter_fields[[name]]
      x <- as.symbol(name) #df[[name]]
      param <- input[[filter[["inputId"]]]]
      cond_expr <- filter[["filterExpr"]](x = x, param = param)
      if (!is.null(cond_expr)) {
        result_expr <<- expr(!!result_expr %>% filter(!!cond_expr))
      }
      invisible()
    })

    result_expr
  })
}

createFilter <- function(data, id, fieldname) {
  UseMethod("createFilter")
}

createFilter.character <- function(data, id, fieldname) {
  list(
    ui = textInput(id, fieldname, ""),
    filterExpr = function(x, param) {
      if (!nzchar(param)) {
        NULL
      } else {
        expr(grepl(!!param, !!x, ignore.case = TRUE, fixed = TRUE))
      }
    }
  )
}

createFilter.numeric <- function(data, id, fieldname) {
  list(
    ui = sliderInput(
      id,
      fieldname,
      min = min(data),
      max = max(data),
      value = range(data)
    ),
    filterExpr = function(x, param) {
      expr(!!x >= !!param[1] & !!x <= !!param[2])
    }
  )
}

createFilter.integer <- createFilter.numeric

createFilter.factor <- function(data, id, fieldname) {
  inputControl <- if (length(levels(data)) > 6) {
    selectInput(id, fieldname, levels(data), character(0), multiple = TRUE)
  } else {
    checkboxGroupInput(id, fieldname, levels(data))
  }

  list(
    ui = inputControl,
    filterExpr = function(x, param) {
      if (length(param) == 0) {
        NULL
      } else {
        expr(!!x %in% !!param)
      }
    }
  )
}

createFilter.POSIXt <- createFilter.numeric
```

### `summarize_module.R`

```{r summarize-module, eval=FALSE}
summarize_ui <- function(id) {
  ns <- NS(id)
  tagList(
    uiOutput(ns("summarize_ui"))
  )
}

summarize_mod <- function(input, output, session, vars, data_expr) {
  output$summarize_ui <- renderUI({
    ns <- session$ns

    tagList(
      selectInput(
        ns("group_by"),
        "Group by",
        choices = vars(),
        multiple = TRUE
      ),
      selectInput(
        ns("operation"),
        "Summary operation",
        c("mean", "sum", "count")
      ),
      selectInput(
        ns("aggregate"),
        "Summary value",
        choices = vars(),
        multiple = TRUE
      )
    )
  })

  reactive({
    result_expr <- data_expr()
    if (length(input$group_by) > 0) {
      result_expr <- expr(!!result_expr %>% group_by(!!!syms(input$group_by)))
    }
    if (length(input$aggregate) > 0) {
      op <- switch(
        input$operation,
        mean = quote(mean),
        sum = quote(sum),
        count = quote(length)
      )
      agg_exprs <- lapply(input$aggregate, function(var) {
        col_name <- deparse(expr((!!sym(input$operation))(!!sym(var))))
        expr(!!col_name := (!!op)(!!sym(var)))
      })
      result_expr <- expr(!!result_expr %>% summarise(!!!agg_exprs))
    }
    result_expr
  })
}
```

